---
title: "Proposal Submission Tracking"
author: "Md Ismail Hossain"
date: "2025-02-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

\newpage

# Data Processing and Merging Overview

This script processes and merges \textbf{proposal data, award data, and college name data} to create a comprehensive dataset for research funding analysis. The process begins by \textbf{loading and cleaning the proposal data}, where cost-related columns are renamed for clarity, and missing statuses are filtered out. Next, the \textbf{award data is loaded and processed similarly}, with cost columns renamed and an additional \texttt{"Proposal \#"} column generated from \texttt{"Award \#"} to facilitate merging. 

After that, the proposal and award datasets are \textbf{merged using a left join on "Proposal \#"}, ensuring that all proposal data is retained while incorporating matching award information. Following this, \textbf{college name data is merged} using \texttt{"PI Unit"} to associate each proposal with its respective college or department. 

Finally, the \texttt{"Created Date"} column is converted into a \textbf{Date-Time format}, and the dataset is \textbf{filtered to include only records from the year 2024}. The resulting dataset provides a \textbf{consolidated view of research proposals, awarded funding, and associated institutions}, making it ready for further analysis or reporting.

# Step 1: Load Required Libraries


```{r}
# Load necessary libraries
library(readr)      # For reading CSV files
library(dplyr)      # For data manipulation
library(readxl)     # For reading Excel files
library(lubridate)  # For handling date-time conversions

```

# Step 2: Load and Process Proposal Data

```{r}
# Read the main Proposal Data file
p_data <- read_csv("C:/Users/mhossa11/OneDrive - University of Wyoming/Projects/Proposal health Project by PI 01312025/Data/Proposal_Data_All_2025_01_31(in).csv")

# Read the Proposal Cost Data file
Proposal_cost_data_all <- read_csv("C:/Users/mhossa11/OneDrive - University of Wyoming/Projects/Proposal health Project by PI 01312025/Data/Proposal_cost_data_all.csv") %>%
  select(-`Project Title`)  # Remove the "Project Title" column

# Rename columns (except "Proposal #") by appending "in Proposal"
colnames(Proposal_cost_data_all)[-1] <- paste0(colnames(Proposal_cost_data_all)[-1], " in Proposal")

# Merge Proposal Data with Proposal Cost Data on "Proposal #"
p_data <- left_join(p_data, Proposal_cost_data_all, by = "Proposal #")

# Select relevant columns for further processing
p_data_selected_columns <- p_data %>%
  select(`Proposal #`, `Project Title`, `Created Date`, PI, `PI Unit`, 
         `History Action`, `History Action Date`, `Status`, 
         `Total Sponsor Costs in Proposal`, `Total Direct Costs in Proposal`, 
         `Total Indirect Costs in Proposal`, `Total Project Cost in Proposal`) %>%
  filter(!is.na(Status))  # Remove rows with missing Status

```


# Step 3: Load and Process Award Data


```{r}
# Read the main Award Data file
a_data <- read_csv("C:/Users/mhossa11/OneDrive - University of Wyoming/Projects/Proposal health Project by PI 01312025/Data/Award_Data_All_2025_01_31(in).csv")

# Read the Award Cost Data file
Award_cost_data_all <- read_csv("C:/Users/mhossa11/OneDrive - University of Wyoming/Projects/Proposal health Project by PI 01312025/Data/Award_cost_data_all.csv") %>%
  select(-`Project Title`)  # Remove "Project Title" column

# Rename columns (except "Award #") by appending "in Award Data"
colnames(Award_cost_data_all)[-1] <- paste0(colnames(Award_cost_data_all)[-1], " in Award Data")

# Merge Award Data with Award Cost Data on "Award #"
a_data <- left_join(a_data, Award_cost_data_all, by = "Award #")

# Select relevant columns for further processing
a_data_selected_columns <- a_data %>%
  select(`Award #`, `Project Title`, PI, `Admin Unit`, `Status`, 
         `Obligated Amount in Award Data`, `Total Expected Amount in Award Data`, 
         `Direct Costs in Award Data`, `expected direct in Award Data`, 
         `expected indirect in Award Data`, `Indirect Costs in Award Data`, 
         `Total Project Cost in Award Data`) %>%
  filter(!is.na(Status))  # Remove rows with missing Status

# Rename "Status" column to "Status_after_award"
colnames(a_data_selected_columns)[5] <- "Status_after_award"

# Generate a "Proposal #" column from "Award #" by replacing "A" with "P"
a_data_selected_columns <- a_data_selected_columns %>%
  mutate(`Proposal #` = gsub("A", "P", `Award #`))

# Perform a left join to merge proposal and award data on "Proposal #"
award_non_award <- p_data_selected_columns %>%
  left_join(a_data_selected_columns, by = "Proposal #")

```


# Step 4: Load and Merge College Name Data

```{r}
# Read the College Name data from an Excel file
college_name <- read_excel("C:/Users/mhossa11/OneDrive - University of Wyoming/Projects/Proposal health Project by PI 01312025/Data/Award External Funding By Dept_Details.xlsx")

# Select relevant columns and remove missing values
college_name_only <- college_name %>%
  select(`College (Subdivision)`, `Department (Organization)`) %>%
  na.omit()

# Rename "Department (Organization)" column to "PI Unit" for merging
colnames(college_name_only)[2] <- "PI Unit"

# Merge college name data with the award/non-award dataset on "PI Unit"
award_non_award_with_college_name <- left_join(college_name_only, award_non_award, by = "PI Unit") %>%
  distinct()  # Remove duplicate rows

```

# Step 5: Convert "Created Date" to Date-Time and Filter for 2024


```{r}
# Convert "Created Date" to Date-Time format
award_non_award_with_college_name <- award_non_award_with_college_name %>%
  mutate(`Created Date` = as.POSIXct(`Created Date`, format = "%m/%d/%Y %H:%M", tz = "UTC"))

# Define the date range for filtering (1st Jan 2024 - 31st Dec 2024)
start_date <- as.POSIXct("2024-01-01 00:00:00", tz = "UTC")
end_date <- as.POSIXct("2024-12-31 23:59:59", tz = "UTC")

# Filter data within the specified date range
filtered_data <- award_non_award_with_college_name %>%
  filter(`Created Date` >= start_date & `Created Date` <= end_date)

# View the filtered data
write.csv(filtered_data,"C:/Users/mhossa11/OneDrive - University of Wyoming/Projects/Proposal health Project by PI 01312025/Output/filtered_data_Jan_2024_Dec_2024_020725.csv")
```

